A comparison of model-based and design-based approaches for spatial data.

### 2021-05-21 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
design-based overview updates

Design-based inference uses characteristics of the sampling design to to estimate parameters of interest Typically, there are few assumptions involved because intervals are derived using the sampling design itself...........

@@

On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimates if the additional assumptions hold. Instead of estimating true but unknown parameters, the goal of model-based inference in the spatial context is often _prediction_ of an unknown quantity. This is a fundamental philosophical difference between sampling-based and model-based approaches. Instead of _estimating_ a fixed unknown mean, we are _predicting_ the value of the mean, a random variable. We know that if we sampled all sites, we would have an exact prediction for the mean of our one realized spatial surface, without any uncertainty. But, the true mean of the spatial process that generated our realized data is still not known, and, in the prediction context, we typically do not care much about what value the mean of the underlying process takes.

>>>>

The design-based approach to inference uses characteristics of the sampling design to estimate parameters of interest. The design-based approach assumes the data, $\mathbf{Y}$, are fixed. Randomness is incorporated through the random selection of sampling units based on probabilities derived using the sampling design itself. Some examples of commonly used sampling designs include independent random sampling (IRS), stratified random sampling, and cluster sampling. Viewing the data as fixed and incorporating randomness through the sampling design yields estimators having very few assumptions. Asymptotic properties of these types of estimators are derived using limiting arguments. Means and totals, for example, are asymptotically Gaussian. @sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.

@@

On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimates if the additional assumptions hold. Instead of estimating true but unknown parameters, the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity. This is a fundamental philosophical difference between sampling-based and model-based approaches. Instead of \emph{estimating} a fixed unknown mean, we are \emph{predicting} the value of the mean, a random variable. We know that if we sampled all sites, we would have an exact prediction for the mean of our one realized spatial surface, without any uncertainty. But, the true mean of the spatial process that generated our realized data is still not known, and, in the prediction context, we typically do not care much about what value the mean of the underlying process takes.

### 2021-05-21 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
design-based overview updates

Design-based inference uses characteristics of the sampling design to to
estimate parameters of interest Typically, there are few assumptions
involved because intervals are derived using the sampling design
itself\ldots\ldots\ldots..

@@

@@

>>>>

The design-based approach to inference uses characteristics of the
sampling design to estimate parameters of interest. The design-based
approach assumes the data, \(\mathbf{Y}\), are fixed. Randomness is
incorporated through the random selection of sampling units based on
probabilities derived using the sampling design itself. Some examples of
commonly used sampling designs include independent random sampling
(IRS), stratified random sampling, and cluster sampling. Viewing the
data as fixed and incorporating randomness through the sampling design
yields estimators having very few assumptions. Asymptotic properties of
these types of estimators are derived using limiting arguments. Means
and totals, for example, are asymptotically Gaussian. Särndal et al.
(2003) and Lohr (2009) provide thorough reviews of the design-based
approach.

@@

\leavevmode\hypertarget{ref-lohr2009sampling}{}%
Lohr, S.L., 2009. Sampling: Design and analysis. Nelson Education.

@@

\leavevmode\hypertarget{ref-sarndal2003model}{}%
Särndal, C.-E., Swensson, B., Wretman, J., 2003. Model assisted survey
sampling. Springer Science \& Business Media.
